15yrsChange

### 2024-04-05 doc/15yrsChange.Rmd
update introduction pt. 5

Understanding the combined effects of multiple nuisance species on forest ecosystem processes requires data on both the forest canopy and understory over an extended time. Here, we use 15 years of detailed tree census from a large forest dynamics plot the mid-Atlantic region of the eastern United States to understand the effects of non-indigenous pests and pathogens and white-tailed deer on the forest carbon balance and regeneration potential.  We test the following hypotheses: (1) canopy tree mortality and associated woody biomass loss are increasing, in large part due to non-indigenous nuisance species, (2) biomass gains from tree growth and recruitment have not kept pace with tree mortality, resulting in a decline in aboveground living biomass (-∆AGB), and (3) deer browsing has decimated the abundance of canopy tree saplings, suggesting that future tree mortality will result in substantial net biomass loss and sustained decrease in the C storage potential of this ecosystem.

>>>>

Understanding the combined effects of multiple nuisance species on forest ecosystem processes requires data on both the forest canopy and understory over an extended time. Here, we use 15 years of detailed tree census data from a large forest dynamics plot in the mid-Atlantic region of the eastern US to understand the effects of non-indigenous pests and pathogens and white-tailed deer on the forest C balance and regeneration potential.  We test the following hypotheses: (1) canopy tree mortality and associated woody biomass loss are increasing, in large part due to non-indigenous nuisance species, (2) biomass gains from tree growth and recruitment have not kept pace with tree mortality, resulting in a decline in aboveground living biomass (-∆AGB), and (3) deer browsing has decimated the abundance of canopy tree saplings, suggesting that future tree mortality will result in substantial net biomass loss and sustained decrease in the C storage potential of this ecosystem.
